\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{epigram}{Aesop, Aesop's Fables}
	\enquote{No act of kindness, no matter how small, is ever wasted.}
\end{epigram}

\noindent
It is to my supervisor, Mark Tame, whom I owe my deepest gratitude. I am greatly indebted for his continuous support, patience and advice, overseeing my research for the past three years. Words have a tendency to debase what their author intends to convey, be it expressions of gratitude, so I will altogether avoid undertaking such an endeavor altogether.

\bigskip
\noindent
I would like to thank Mueletshedzi for her Herculean resolve of proof-reading an early version of the front matter of this document, companionship and all the coffee breaks, Olivia and Jacques for their charitable acts of kindness, Hjalmar for translating the abstract of this document, and listening to my ramblings down in the lab, Andre for diagnosing an interference filter used in one of the experiments here, Mr Botha for technical support down in the lab, and Khilly for being a great companion.

\bigskip
\noindent
I would also like to extend my thanks to Taariq Surtee and Barry Dwolatzky at the University of the Witwatersrand and Ismail Akhalwaya at IBM Research Africa for access to the IBM processors through the Q Network and African Research Universities Alliance. Half of this thesis would not have come to fruition if it were not for them. This research was supported by the South African National Research Foundation, the South African Council for Scientific and Industrial Research, and the South African Research Chair Initiative of the Department of Science and Technology and National Research Foundation.

\bigskip
\noindent
Most importantly, I would like to especially thank my family for their support and encouragement to pursue whatever it was I happened to be interested in, and granting me an opportunity to do so.
